---
title: 'Status of the Yellowtail Rockfish Stock off the U.S. West Coast North of 40°10$\prime$ in 2025'
author:
  - name: 'Kiva L. Oken'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        postal-code: '98112-2097'
  - name: 'Ian G. Taylor'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        postal-code: '98112-2097'
  - name: 'Megan L. Feddern'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        postal-code: '98112-2097'
  - name: 'Alison D. Whitman'
    affiliations:
      - name: 'Oregon Department of Fish and Wildlife'
        address: '2040 Southeast Marine Science Drive'
        city: 'Newport, OR'
        postal-code: '97365'
  - name: 'Fabio P. Caltabellotta'
    affiliations:
      - name: 'Washington Department of Fish and Wildlife'
        address: '600 Capital Way North'
        city: 'Olympia, WA'
        postal-code: '98501'
date: today
lang: en 
keep-tex: true 
cover: 
pdf-engine: lualatex
format: 
  pdf: 
    documentclass: scrartcl
    number-sections: true
    template-partials:
      - 'support_files/before-body.tex'
      - 'support_files/_titlepage.tex'
    include-in-header:
      - 'support_files/in-header.tex'
      - 'support_files/pfmc.tex'
    toc: false
    lof: false
    lot: false
    titlepage-geometry: 
      - top=2in
      - bottom=1in
      - right=1in
      - left=1in
    link-citations: true
    fig-pos: 'H'
output-file: 'Yellowtail_rockfish_SAR_2025' 
params:
   office: 'NWFSC'
   species: 'Yellowtail rockfish'
   spp_latin: 'Sebastes flavidus'
   region: 'U.S. West Coast north of 40°10$\prime$'
bibliography: 
  -  asar_references.bib 
  -  yellowtail.bib
csl: 'canadian-journal-of-fisheries-and-aquatic-sciences.csl'
---

\pagenumbering{roman}
\setcounter{page}{1}

\renewcommand{\thetable}{\roman{table}}
\renewcommand{\thefigure}{\roman{figure}}

```{r}
#| label: 'read_base_model'
#| eval: true
#| echo: false
#| message: false
#| warning: false

library(here)
library(dplyr)

source(here::here('Rscripts/table_sens.R'))
source(here::here('Rscripts/table_decision.R'))

# TODO list of things to update when changing the base model:
# - [ ] run Rscripts/make_r4ss_figs_tables.R
# - [ ] run Rscripts/run_diagnostics.R
# - [ ] run Rscripts/sensitivities.R
# - [ ] run Rscript/CatchTable.R if catches changed
# - [ ] update any hardwired values in the text (or make them dynamic)
#   - [ ] text on 2024 discards "inadvertently left out of the base model"
#   - [ ] if changing selectivity, try freeing up triennial peak selectivity (fixed at 55cm)
#   - [ ] if changing selectivity, try freeing up hake peak selectivity (fixed at 55cm)
base_mod <- '5.09_no_extra_se'
mod_in <- r4ss::SS_read(here('model_runs', base_mod))
mod_out <- r4ss::SS_output(here('model_runs', base_mod), verbose = FALSE, printstats = FALSE)

```

{{< pagebreak >}} 

```{r, results='asis'}
#| label: 'disclaimer'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('00a_disclaimer.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

{{< pagebreak >}} 

Please cite this publication as: 

Oken, K.L., I.G. Taylor, M.L. Feddern, A.D. Whitman, F.P. Caltabellotta. Status of the yellowtail rockfish stock off the U.S. West Coast north of 40°10$\prime$ in 2025. Pacific Fishery Management Council, Portland, Oregon. [XX] p.

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'load_tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('002_load_tables.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

```{=latex}
{
\hypersetup{linkcolor=.}
\setcounter{tocdepth}{3}
\tableofcontents
}
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'executive_summary'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('01_executive_summary.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\setcounter{section}{0}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}


```{r, results='asis'}
#| label: 'introduction'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('02_introduction.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'data'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('03_data.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'assessment-configuration'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04a_model_summary.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: 'assessment-sensitivity'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04b_model_results.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: 'assessment-benchmarks'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04c_model_diagnostics.qmd', quiet = TRUE) 
cat(a, sep = '\n') 
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'discussion'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('05_management.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'acknowledgments'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('06_acknowledgments.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'references'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('07_references.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('08_tables.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'figures'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('09_figures.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

\appendix

```{r, results='asis'}
#| label: 'appendix'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('11_appendix.qmd', quiet = TRUE)
cat(a, sep = '\n')
```
