---
title: "Data summary"
date: today
format: 
  html:
    toc: true
    output-file: "index"
editor: visual
---

```{r opts, echo = FALSE, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(here)
library(dplyr)
library(ggplot2)
theme_set(theme_classic())
```

This documents data streams for use in the Northern Yellowtail Rockfish stock assessment. These data are NOT final and have NOT been fully QA/QC'ed! They are here for communication with the STAT and other collaborators. It is not official communication of the National Oceanic and Atmospheric Administration or the United States Department of Commerce.

## Landings

### Commercial landings
Some summary figures:
```{r pacfin-catch, cache=TRUE}
load(here('data/confidential/commercial/PacFIN.YTRK.CompFT.09.Oct.2024.RData'))
# I have checked and LANDING_YEAR = PACFIN_YEAR for all rows.

yt_n_catch_pacfin <- catch.pacfin |>
  filter(
    AGENCY_CODE == 'W' | AGENCY_CODE == 'O' | COUNTY_NAME == 'DEL NORTE' | COUNTY_NAME == 'HUMBOLDT',
    LANDING_YEAR < 2024
  )

# OR landings time series
or_comm_catch <- read.csv(here("Data/Confidential/commercial/Oregon Commercial landings_433_2023.csv")) |>
  tibble::as_tibble() |>
  group_by(YEAR) |>
  summarise(catch_mt = sum(TOTAL))

comm_draft_landings <- yt_n_catch_pacfin |> 
  filter(AGENCY_CODE != 'O') |> 
  group_by(PACFIN_YEAR, AGENCY_CODE)|>
  summarise(catch_mt = sum(LANDED_WEIGHT_MTONS)) |>
  rename(YEAR = PACFIN_YEAR) |>
  bind_rows(
    mutate(or_comm_catch, AGENCY_CODE = 'O')
  )|> 
  group_by(YEAR, AGENCY_CODE) |>
  summarise(landings_mt = sum(catch_mt)) |>
  filter(YEAR >= 1981) 

# 2017 assessment landings
comm2017 <- read.csv(here("Data/Raw_not_confidential/commercial_catch_2017.csv")) |> 
  as_tibble() |>
  tidyr::replace_na(list(Mirick.reconstructed.POP1.URCK = 0)) |>
  mutate(O = OR.Recent..MT. + Mirick.reconstructed.POP1.URCK + OR.Hist..MT.,
         W = as.numeric(stringr::str_remove(WA, ','))) |> 
  select(Year, W, C = CA.North..MT., O) |>
  filter(Year >= 1981) |>
  tidyr::pivot_longer(-Year, names_to = 'AGENCY_CODE', values_to = 'landings_mt') |>
  mutate(model = '2017')

# landings by state
comm_draft_landings |>
  ggplot() +
  geom_col(aes(x = YEAR, y = landings_mt, fill = AGENCY_CODE)) +
  ggtitle('Commercial landings through 2023 by state') +
  viridis::scale_fill_viridis(discrete = TRUE)

# landings by gear type
yt_n_catch_pacfin |>
  group_by(PACFIN_GROUP_GEAR_CODE, LANDING_YEAR) |>
  summarise(landings_mt = sum(LANDED_WEIGHT_MTONS)) |>
  ggplot() +
  geom_col(aes(x = LANDING_YEAR, y = landings_mt, fill = factor(PACFIN_GROUP_GEAR_CODE, levels = c('TWL', 'MSC', 'NET', 'POT', 'TLS', 'TWS', 'HKL')))) +
  ggtitle('Commercial landings through 2023 by gear') +
  labs(fill = 'PacFIN gear group', x = 'YEAR') +
  viridis::scale_fill_viridis(discrete = TRUE)

# total catch by gear over time with midwater trawl separated
yt_n_catch_pacfin |>
  mutate(geargroup2 = ifelse(PACFIN_GEAR_CODE == "MDT", yes = "MDT", no = PACFIN_GROUP_GEAR_CODE)) |> # split midwater trawl as MDT
  group_by(geargroup2, LANDING_YEAR) |>
  summarise(catch_mt = sum(LANDED_WEIGHT_MTONS)) |>
  ggplot() +
  geom_col(aes(x = LANDING_YEAR, y = catch_mt, fill = factor(geargroup2, levels = c("TWL", "MDT", "MSC", "NET", "POT", "TLS", "TWS", "HKL")))) +
  ggtitle("Commercial landings through 2023") +
  labs(fill = "PacFIN gear group\n(MDT = midwater trawl)", x = 'YEAR') #+
  # viridis::scale_fill_viridis(discrete = TRUE) # colors were hard to see in viridis

# compare 2017 landings to current landings, by state
comm_draft_landings |>
  rename(Year = YEAR) |>
  mutate(model = 'current') |>
  bind_rows(comm2017) |> 
  ggplot() +
  geom_line(aes(x = Year, y = landings_mt, col = AGENCY_CODE, linetype = model), linewidth = 1) +
  viridis::scale_color_viridis(discrete = TRUE) +
  labs(title = 'Comparing commercial landings in 2017 model vs current', x = 'YEAR')
```
Table of commercial landings by state, 1981 onward.

```{r comm-catch-tbl}
comm_draft_landings |>
  mutate(State = case_when(AGENCY_CODE == 'C' ~ 'CA',
                           AGENCY_CODE == 'O' ~ 'OR',
                           AGENCY_CODE == 'W' ~ 'WA')) |>
  select(-AGENCY_CODE) |>
  tidyr::pivot_wider(names_from = State, values_from = landings_mt) |>
  knitr::kable(align = 'l', digits = 1)
```

WDFW has alerted the STAT these do not include all tribal catches in recent years. 


Note this is for catch landed into Del Norte and Humboldt counties only.

### At-Sea landings

```{r ashop-catch}
ashop_catch <- readxl::read_excel(
  here("Data/Confidential/ASHOP/Oken_YLT_Catch data_1976-2023_102824_ASHOP.xlsx"), 
  sheet = 'Catch Summary Table')
ashop_catch |>
  mutate(catch_mt = EXPANDED_SumOfEXTRAPOLATED_2SECTOR_WEIGHT_KG / 1000) |>
  select(YEAR, catch_mt) |>
  knitr::kable(align = 'l', digits = 1)
```


### Recreational landings

```{r rec-catch-modern}
wa_modern <- read.csv(here('Data/Raw_not_confidential/RecFIN_WA_catch_to_2023.csv')) |>
  filter(RECFIN_WATER_AREA_NAME != 'Canada', RECFIN_YEAR <= 2023) |>
  tibble::as_tibble() |>
  group_by(RECFIN_YEAR) |>
  summarise(Dead_Catch = sum(SUM_TOTAL_MORTALITY_MT)) |>
  mutate(State = 'WA (mt)')

wa_historical <- read.csv(here('Data/Raw_not_confidential/WA_historical_rec.csv')) |>
  tibble::as_tibble() |>
  group_by(RECFIN_YEAR) |>
  summarise(Dead_Catch = sum(RETAINED_NUM)) |>
  mutate(State = 'WA (numbers)')

or_rec_catch <- read.csv(here("Data/Confidential/Rec/Oregon Recreational landings_433_2023.csv")) |>
  tibble::as_tibble() |>
  mutate(State = 'OR (mt)')

ca_modern <- read.csv(here('Data/Raw_not_confidential/RecFIN_CA_catch_to_2023.csv')) |>
  filter(grepl("Redwood", DISTRICT_NAME)) |>
  tibble::as_tibble() |>
  group_by(RECFIN_YEAR) |>
  summarise(Dead_Catch = sum(SUM_TOTAL_MORTALITY_MT)) |>
  mutate(State = 'CA (mt)')

or_rec_catch |>
  select(RECFIN_YEAR = Year, Dead_Catch = Total_MT, State) |>
  bind_rows(wa_modern, wa_historical, ca_modern) |>
  tidyr::pivot_wider(names_from = 'State', values_from = 'Dead_Catch', values_fill = 0) |>
  rename(YEAR = 'RECFIN_YEAR') |>
  arrange(YEAR) |>
  knitr::kable(align = 'l', digits = 1)
```

Outstanding issues for WA:

1.  There are three values for `RECFIN_WATER_AREA_NAME`: Estuary, Ocean, and Canada. Which should be included? Above table excludes Canada.
2.  2023 has a number of instances where `RECFIN_WEEK` is zero, and one instance where is it missing. In the instance where it is missing, there is no estimate of catch in mt.
3.  Will likely convert the numbers in the historical sport landings to weight in mt. On the to do list.

Outstanding issues for CA: 

1. 2020 and 2021 proxy catches and un-ID'ed rockfish apportionment are missing.
2. MRFSS data (1981-2004) does not have an obvious way to determine whether catches should be assigned to the northern or southern stock.


## Comp Data

### Commercial

Initial length sample sizes after running `PacFIN.Utilities::cleanPacFIN()`:

```{r comm-bio, cache=TRUE}
load(here('Data/Confidential/Commercial/PacFIN.YTRK.bds.09.Oct.2024.RData'))
bds_clean <- PacFIN.Utilities::cleanPacFIN(bds.pacfin, verbose = FALSE) 
bds_clean |>
  filter(state == 'WA' | state == 'OR' | PACFIN_GROUP_PORT_CODE == 'CCA' |
           PACFIN_GROUP_PORT_CODE == 'ERA', geargroup != 'XXX') |> 
  filter(year < 2024) |>
  group_by(state, year) |>
  summarise(n_length = sum(!is.na(length))) |>
  tidyr::pivot_wider(names_from = state, values_from = n_length, values_fill = 0) |>
  rowwise() |>
  mutate(total = OR + WA + CA) |>
  arrange(year) |>
  knitr::kable(align = 'l', digits = 1)

```

Age sample sizes:
```{r comm-ages}
bds_clean |>
  filter(state == 'WA' | state == 'OR' | PACFIN_GROUP_PORT_CODE == 'CCA' |
           PACFIN_GROUP_PORT_CODE == 'ERA') |> 
  filter(year < 2024) |>
  group_by(state, year) |>
  summarise(n_length = sum(!is.na(Age))) |>
  tidyr::pivot_wider(names_from = state, values_from = n_length, values_fill = 0) |>
  rowwise() |>
  mutate(total = OR + WA + CA) |>
  arrange(year) |>
  knitr::kable(align = 'l', digits = 1)
```

Note that California lengths are not yet available for 2023.

While running `PacFIN.Utilities::cleanPacFIN()`, it noted that 20 Washington ages in 1991 were double read, the second reader was not recorded, the first and second ages differed, and no final age was determined.

#### Comparing midwater vs bottom trawl lengths and ages

The length and age distribution associated with the midwater trawl (MDT) code in PacFIN is is very similar to bottom trawl catch (everything else in the TWL gear group). Note that the sample sizes for ages in 2022-2024 are lower than other years, especially midwater in 2022 which only has 14 ages.

```{r pacfin-midwater-vs-bottom-figs}
# create separate geargroup2 column which separates midwater from bottom trawl
# see note at 
# https://github.com/pfmc-assessments/PacFIN.Utilities/issues/69#issuecomment-2499162809
bds_clean <- bds_clean |>
  dplyr::mutate(geargroup2 = ifelse(GRID == "MDT", yes = "MDT", no = geargroup))

# length distributions for bottom trawl vs midwater trawl in recent years
bds_clean |>
  filter(geargroup == "TWL", SAMPLE_YEAR >= 2011) |> # exclude all non-trawl gear and pre-catch-shares years
  ggplot(aes(
    x = lengthcm,
    y = forcats::fct_rev(as.factor(SAMPLE_YEAR)),
    color = as.factor(geargroup2)
    #fill = as.factor(geargroup2)
  )) +
  ggridges::geom_density_ridges(alpha = 0.2) +
  labs(
    x = "Length (cm)",
    y = "Year",
  ) + 
  scale_color_discrete(
    name = "Gear",
    breaks = c("MDT", "TWL"),
    labels = c("Midwater trawl", "Bottom trawl")
  )

# age comps from midwater vs bottom trawl
bds_clean |>
  filter(geargroup == "TWL", !is.na(Age),SAMPLE_YEAR >= 2011) |>
  ggplot(aes(
    x = Age,
    y = forcats::fct_rev(as.factor(SAMPLE_YEAR)),
    color = as.factor(geargroup2)
  )) +
  ggridges::geom_density_ridges(
    alpha = 0.2,
    stat = "binline",
    binwidth = 1
  ) +
  labs(
    x = "Age",
    y = "Year",
  ) +
  scale_color_discrete(
    name = "Gear",
    breaks = c("MDT", "TWL"),
    labels = c("Midwater trawl", "Bottom trawl")
  )

```


### Recreational

VERY tentative length sample sizes. I anticipate this is biased high as I have done no filtering.

```{r recfin-bio}
# recent length data
rec_bio <- read.csv(here("Data/Confidential/rec/RecFIN_Lengths.csv")) |>
  tibble::as_tibble() |>
  filter(STATE_NAME == "OREGON" | STATE_NAME == "WASHINGTON" | grepl("REDWOOD", RECFIN_PORT_NAME))

# or mrfss data
or_mrfss <- read.csv(here("Data/Confidential/rec/MRFSS_BIO_2025 CYCLE_433.csv")) |>
  tibble::as_tibble()

or_mrfss_smry <- or_mrfss |> 
  filter(Length_Flag == 'measured') |>
  group_by(Year) |> 
  summarise(n_length = n()) |>
  mutate(STATE_NAME = 'OR (MRFSS)') |>
  tidyr::pivot_wider(names_from = STATE_NAME, values_from = n_length)

# ca/wa mrfss data
ca_wa_mrfss <- read.csv(here("Data/Confidential/rec/MRFSS_lengths_CA_WA.csv")) |>
  as_tibble() |> 
  filter(ST_NAME == 'Washington' | (ST_NAME == 'California' & (CNTY == 15 | CNTY == 23))) |> # 15 = Del Norte, 23 = Humboldt, FIPS codes
  filter(T_LEN %% 1 == 0 | LNGTH %% 1 == 0) # I believe this will filter to measured lengths

ca_wa_mrfss_smry <- ca_wa_mrfss |>
  group_by(YEAR, ST_NAME) |>
  summarise(n_length = n()) |>
  ungroup() |>
  mutate(state = case_when(ST_NAME == 'California' ~ 'CA (MRFSS)',
                           ST_NAME == 'Washington' ~ 'WA (MRFSS)')) |>
  tidyr::pivot_wider(names_from = state, values_from = n_length, id_cols = YEAR, values_fill = 0) |>
  rename(Year = YEAR)

rec_bio |>
  ggplot() +
  geom_density(aes(x = RECFIN_LENGTH_MM, col = STATE_NAME, fill = IS_RETAINED), 
               linewidth = 1, alpha = 0.2) +
  scale_fill_manual(values = c('red', 'blue')) +
  labs('RecFIN length compositions by state and disposition') +
  viridis::scale_color_viridis(discrete = TRUE) 

# Note: 1993 uses total length. 1994 onward uses fork length.
rec_bio |>
  filter(!is.na(RECFIN_LENGTH_MM)) |> 
  mutate(state = case_when(STATE_NAME == 'WASHINGTON' ~'WA',
                           STATE_NAME == 'OREGON' ~ 'OR',
                           STATE_NAME == 'CALIFORNIA' ~ 'CA')) |>
  group_by(RECFIN_YEAR, state, IS_RETAINED) |>
  summarize(n_length = n()) |>
  rename(Year = RECFIN_YEAR) |> 
  tidyr::pivot_wider(names_from = c(state, IS_RETAINED), values_from = n_length, 
                     values_fill = 0) |> 
  full_join(or_mrfss_smry) |>
  full_join(ca_wa_mrfss_smry) |>
  arrange(Year) |> 
  mutate(across(everything(), ~tidyr::replace_na(.x, 0))) |>
  knitr::kable(align = 'l')
```

Retained fish tend to be larger than released fish. However, there are very few released fish. Washington has no measured released fish in RecFIN. Between Oregon and California, 
`r filter(rec_bio, STATE_NAME != 'WASHINGTON') %>% {100 * with(., sum(IS_RETAINED == 'RETAINED'))/nrow(.)} |> round(1)`\% of lengths are for retained fish. Some fraction of those released fish is assumed to have survived, which would skew the ratio even more towards retained fish.

Options:

1. Exclude released fish from comps
2. Lump released and retained fish together
3. Model a retention curve

The fish in Washington also look slightly larger than those in Oregon and California. This difference is more pronounced than it is in survey data, indicating a possible selectivity effect.

I am unsure how to filter MRFSS data to measured lengths for WA and CA. The counts in the table above filters lengths to whole numbers (mm).

### At-Sea
```{r ashop-bio}
ashop_lengths_old <- readxl::read_excel(
  here("Data/Confidential/ASHOP/Oken_YLT_Length data_1976-2023_102824_ASHOP.xlsx"), 
  sheet = "YLT_Length data 1976-1989")
ashop_lengths_new <- suppressWarnings(
  readxl::read_excel(here("Data/Confidential/ASHOP/Oken_YLT_Length data_1976-2023_102824_ASHOP.xlsx"), 
                     sheet = "YLT_Length data1990-2023") # warning is about converting text to numeric
)

bind_rows(ashop_lengths_old, ashop_lengths_new) |>
  group_by(YEAR) |>
  summarise(n_length = sum(FREQUENCY)) |>
  knitr::kable(align = 'l')
```

Figure of length frequencies in at-sea hake fishery for prioritizing which years to age.

```{r ashop-bio-fig}
ashop_lengths_new |> 
  tidyr::uncount(FREQUENCY) |>
  left_join(ashop_catch, by = 'YEAR') |> 
  mutate(catch_mt = EXPANDED_SumOfEXTRAPOLATED_2SECTOR_NUMBER / 1000) |>
  ggplot() +
  ggridges::geom_density_ridges(aes(x = LENGTH, y = YEAR, group = YEAR, 
                                    fill = catch_mt), alpha = 0.5) +
  xlim(30, 65) +
  viridis::scale_fill_viridis()
```

## Survey data

STAT is working on it!

```{r survey, eval=FALSE}
yt_survey_bio <- nwfscSurvey::pull_bio(common_name = 'yellowtail rockfish', survey = 'NWFSC.Combo') 

yt_n_survey_bio <- filter(yt_survey_bio, Latitude_dd > 40 + 1/6)

yt_survey_bio |>
  filter(!is.na(Otosag_id),
         is.na(Age_years)) |>
  group_by(Year) |>
  summarise(n())

# clear break point in selectivity between age 5 and 6.

filter(yt_n_survey_bio, Age_years <= 20, Age_years >= 6) |>
ggplot() +
  geom_point(aes(x = Year, y = Length_cm, col = Age_years), alpha = 0.2) +
  stat_smooth(aes(x = Year, y = Length_cm, group = Age_years, col = Age_years), se = FALSE)

filter(yt_survey_bio, Age_years <= 20, Age_years >= 6) |>
ggplot() +
  geom_point(aes(x = Latitude_dd, y = Length_cm, col = Age_years), alpha = 0.2) +
  stat_smooth(aes(x = Latitude_dd, y = Length_cm, group = Age_years, col = Age_years), se = FALSE, method = 'lm')
# possible decrease in length at age of older fish in last 4 years of data. TBD with 8 more years!

yt_survey_bio |>
  ggplot(aes(x = Latitude_dd, y = Age_years)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE)
# Latitudinal age structure: below 43 degrees, avg age ~10, above 45 degrees, avg age ~15

ggplot(yt_n_survey_bio) +
  geom_point(aes(x = Age_years, y = Length_cm, col = Sex), alpha = 0.1)
# Large fish female, old fish male prevails, but not as notable as with canary.

yt_n_survey_bio |>
  group_by(Age_years) |>
  summarise(pct_f = sum(Sex == 'F')/n(),
            se = sqrt(pct_f * (1-pct_f) / n())) |>
  ggplot() +
  geom_point(aes(x = Age_years, y = pct_f)) +
  geom_errorbar(aes(x = Age_years, ymin = pct_f - se, ymax = pct_f + se))

laa_by_lat_gam <- lm(Length_cm ~ I(Latitude_dd-mean(Latitude_dd))*factor(Age_years), 
                                   data = yt_n_survey_bio, 
                      subset = Age_years < 20 & Age_years >= 6)

laa_by_age <- lm(Length_cm ~ factor(Age_years), data = yt_n_survey_bio, 
                      subset = Age_years < 20 & Age_years >= 6)

```

