---
title: "Data summary"
date: today
format: 
  html:
    toc: true
    output-file: "index"
editor: visual
---

```{r opts, echo = FALSE, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(here)
library(dplyr)
library(ggplot2)
theme_set(theme_gray())
```

This documents data streams for use in the Northern Yellowtail Rockfish stock assessment. These data are NOT final and have NOT been fully QA/QC'ed! They are here for communication with the STAT and other collaborators. It is not official communication of the National Oceanic and Atmospheric Administration or the United States Department of Commerce.

## Commercial landings

```{r pacfin-catch, cache=TRUE}
load(here('data/confidential/PacFIN.YTRK.CompFT.09.Oct.2024.RData'))

yt_n_catch <- catch.pacfin |>
  filter(
    AGENCY_CODE == 'W' | AGENCY_CODE == 'O' | COUNTY_NAME == 'DEL NORTE' | COUNTY_NAME == 'HUMBOLDT',
    LANDING_YEAR < 2024
  )

yt_n_catch |>
  group_by(AGENCY_CODE, LANDING_YEAR) |>
  summarise(catch_mt = sum(LANDED_WEIGHT_MTONS)) |>
  ggplot() +
  geom_col(aes(x = LANDING_YEAR, y = catch_mt, fill = AGENCY_CODE)) +
  ggtitle('Commercial landings through 2023')

yt_n_catch |>
  group_by(PACFIN_GROUP_GEAR_CODE, LANDING_YEAR) |>
  summarise(catch_mt = sum(LANDED_WEIGHT_MTONS)) |>
  ggplot() +
  geom_col(aes(x = LANDING_YEAR, y = catch_mt, fill = factor(PACFIN_GROUP_GEAR_CODE, levels = c('TWL', 'MSC', 'NET', 'POT', 'TLS', 'TWS', 'HKL')))) +
  ggtitle('Commercial landings through 2023') +
  labs(fill = 'PacFIN gear group') +
  viridis::scale_fill_viridis(discrete = TRUE)
```

### Washington

```{r wa-comm-catch}
filter(yt_n_catch, AGENCY_CODE == 'W') |>
  group_by(LANDING_YEAR) |>
  summarise(catch_mt = sum(LANDED_WEIGHT_MTONS)) |>
  knitr::kable(align = 'l', digits = 1)
```

WDFW has alerted the STAT these do not include all tribal catches in recent years. 

### Oregon

The following values were provided directly to the STAT:
```{r or-comm-catch}
or_comm_catch <- read.csv(here("Data/Confidential/Oregon Commercial landings_433_2023.csv")) |>
  tibble::as_tibble()
or_comm_catch |> 
  group_by(YEAR) |>
  summarise(catch_mt = sum(TOTAL)) |>
  knitr::kable(align = 'l', digits = 1)
```


### California

```{r ca-comm-catch}
filter(yt_n_catch, AGENCY_CODE == 'C') |>
  group_by(LANDING_YEAR) |>
  summarise(catch_mt = sum(LANDED_WEIGHT_MTONS))  |>
  knitr::kable(align = 'l', digits = 1)
```

Note this is for catch landed into Del Norte and Humboldt counties only.

## At-Sea landings

```{r ashop-catch}
ashop_catch <- readxl::read_excel(
  here("Data/Confidential/Oken_YLT_Catch data_1976-2023_102824_ASHOP.xlsx"), 
  sheet = 'Catch Summary Table')
ashop_catch |>
  mutate(catch_mt = EXPANDED_SumOfEXTRAPOLATED_2SECTOR_WEIGHT_KG / 1000) |>
  select(YEAR, catch_mt) |>
  knitr::kable(align = 'l', digits = 1)
```


## Recreational landings

### Washington

Modern catches:

```{r wa-rec-catch-modern}
wa_modern <- read.csv(here('Data/Raw_not_confidential/RecFIN_WA_catch_to_2023.csv')) |>
  filter(RECFIN_WATER_AREA_NAME != 'Canada', RECFIN_YEAR < 2023) |>
  tibble::as_tibble()
wa_modern |> group_by(RECFIN_YEAR) |>
  summarise(Dead_Catch_mt = sum(SUM_TOTAL_MORTALITY_MT))  |>
  knitr::kable(align = 'l', digits = 1)
```

Questions:

1.  There are three values for `RECFIN_WATER_AREA_NAME`: Estuary, Ocean, and Canada. Which should be included? Above table excludes Canada.
2.  2023 has a number of instances where `RECFIN_WEEK` is zero, and one instance where is it missing. In the instance where it is missing, there is no estimate of catch in mt.

Historical catches:

```{r wa-rec-catch-historical}
wa_historical <- read.csv(here('Data/Raw_not_confidential/WA_historical_rec.csv')) |>
  tibble::as_tibble()

wa_historical |>
  group_by(RECFIN_YEAR) |>
  summarise(Dead_Catch_num = sum(RETAINED_NUM)) |>
  knitr::kable(align = 'l', digits = 1)
```

### Oregon

The following values were provided directly to the STAT:
```{r or-rec-catch}
or_rec_catch <- read.csv(here("Data/Confidential/Oregon Recreational landings_433_2023.csv")) |>
  tibble::as_tibble()
or_rec_catch |>
  select(Year, Total_MT) |>
  knitr::kable(align = 'l', digits = 1)
```


### California

```{r ca-rec-catch}
ca_modern <- read.csv(here('Data/Raw_not_confidential/RecFIN_CA_catch_to_2023.csv')) |>
  filter(grepl("Redwood", DISTRICT_NAME)) |>
  tibble::as_tibble()
ca_modern |>
  group_by(RECFIN_YEAR) |>
  summarise(Dead_Catch_mt = sum(SUM_TOTAL_MORTALITY_MT))  |>
  knitr::kable(align = 'l', digits = 1)
```

Note that 2020 proxy catches are missing, and we will need historical recreational catches that are not on RecFIN.

MRFSS data (1981-2004) does not have an obvious way to determine whether catches should be assigned to the northern or southern stock.

## Length Data

### Commercial

Initial sample sizes after running `PacFIN.Utilities::cleanPacFIN()`:

```{r comm-bio, cache=TRUE}
load(here('Data/Confidential/PacFIN.YTRK.bds.09.Oct.2024.RData'))
bds_clean <- PacFIN.Utilities::cleanPacFIN(bds.pacfin, verbose = FALSE) 
bds_clean |>
  filter(state == 'WA' | state == 'OR' | PACFIN_GROUP_PORT_CODE == 'CCA' |
           PACFIN_GROUP_PORT_CODE == 'ERA', geargroup != 'XXX') |> 
  filter(year < 2024) |>
  group_by(state, year) |>
  summarise(n_length = sum(!is.na(length))) |>
  tidyr::pivot_wider(names_from = state, values_from = n_length, values_fill = 0) |>
  arrange(year) |>
  knitr::kable(align = 'l', digits = 1)

```

Note that California lengths are not yet available for 2023.

While running `PacFIN.Utilities::cleanPacFIN()`, it noted that 20 Washington ages in 1991 were double read, the second reader was not recorded, the first and second ages differed, and no final age was determined.

### Recreational

VERY tentative length sample sizes. I anticipate this is biased high as I have done no filtering.

```{r rec-bio}
rec_bio <- read.csv(here("Data/Confidential/RecFIN_Lengths.csv")) |>
  tibble::as_tibble() |>
  filter(STATE_NAME == "OREGON" | STATE_NAME == "WASHINGTON" | grepl("REDWOOD", RECFIN_PORT_NAME))

or_mrfss <- read.csv(here("Data/Confidential/MRFSS_BIO_2025 CYCLE_433.csv")) |>
  tibble::as_tibble()

or_mrfss_smry <- or_mrfss |> 
  filter(Length_Flag == 'measured') |>
  group_by(Year) |> 
  summarise(n_length = n()) |>
  mutate(STATE_NAME = 'OREGON')

filter(rec_bio, STATE_NAME == "OREGON", RECFIN_YEAR %in% or_mrfss_smry$Year) |>
  group_by(RECFIN_YEAR) |>
  filter(!is.na(RECFIN_LENGTH_MM)) |>
  arrange(RECFIN_YEAR) |>
  summarise(n_length = n())

rec_bio |>
  filter(!is.na(RECFIN_LENGTH_MM), STATE_NAME != 'OREGON' | !(RECFIN_YEAR %in% or_mrfss_smry$Year),
         RECFIN_LENGTH_TYPE == TRUE) |>
  group_by(RECFIN_YEAR, STATE_NAME) |>
  summarize(n_length = n()) |>
  rename(Year = RECFIN_YEAR) |>
  bind_rows(or_mrfss_smry) |> 
  tidyr::pivot_wider(names_from = STATE_NAME, values_from = n_length, values_fill = 0) |> 
  arrange(Year) |>
  knitr::kable(align = 'l')
```
From Oregon MRFSS data (provided directly to the STAT, covering `r min(or_mrfss_smry$Year)` to `r max(or_mrfss_smry$Year)`) this only includes counts of samples with `Length_Flag == "measured"`. I was not sure of recommendations for how to filter the data.

### At-Sea
```{r ashop-bio}
ashop_lengths_old <- readxl::read_excel(
  here("Data/Confidential/Oken_YLT_Length data_1976-2023_102824_ASHOP.xlsx"), 
  sheet = "YLT_Length data 1976-1989")
ashop_lengths_new <- suppressWarnings(
  readxl::read_excel(here("Data/Confidential/Oken_YLT_Length data_1976-2023_102824_ASHOP.xlsx"), 
                     sheet = "YLT_Length data1990-2023") # warning is about converting text to numeric
)

bind_rows(ashop_lengths_old, ashop_lengths_new) |>
  group_by(YEAR) |>
  summarise(n_length = sum(FREQUENCY)) |>
  knitr::kable(align = 'l')
```

Figure of length frequencies in at-sea hake fishery for prioritizing which years to age.

```{r ashop-bio-fig}
ashop_lengths_new |> 
  tidyr::uncount(FREQUENCY) |>
  left_join(ashop_catch, by = 'YEAR') |> 
  mutate(catch_mt = EXPANDED_SumOfEXTRAPOLATED_2SECTOR_NUMBER / 1000) |>
  ggplot() +
  ggridges::geom_density_ridges(aes(x = LENGTH, y = YEAR, group = YEAR, 
                                    fill = catch_mt), alpha = 0.5) +
  xlim(30, 65) +
  viridis::scale_fill_viridis()
```

## Survey data

STAT is working on it!

```{r survey, eval=FALSE}
yt_survey_bio <- nwfscSurvey::pull_bio(common_name = 'yellowtail rockfish', survey = 'NWFSC.Combo') 

yt_n_survey_bio <- filter(yt_survey_bio, Latitude_dd > 40 + 1/6)

yt_survey_bio |>
  filter(!is.na(Otosag_id),
         is.na(Age_years)) |>
  group_by(Year) |>
  summarise(n())

# clear break point in selectivity between age 5 and 6.

filter(yt_n_survey_bio, Age_years <= 20, Age_years >= 6) |>
ggplot() +
  geom_point(aes(x = Year, y = Length_cm, col = Age_years), alpha = 0.2) +
  stat_smooth(aes(x = Year, y = Length_cm, group = Age_years, col = Age_years), se = FALSE)

filter(yt_survey_bio, Age_years <= 20, Age_years >= 6) |>
ggplot() +
  geom_point(aes(x = Latitude_dd, y = Length_cm, col = Age_years), alpha = 0.2) +
  stat_smooth(aes(x = Latitude_dd, y = Length_cm, group = Age_years, col = Age_years), se = FALSE)
# possible decrease in length at age of older fish in last 4 years of data. TBD with 8 more years!

yt_survey_bio |>
  ggplot(aes(x = Latitude_dd, y = Age_years)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE)
# Latitudinal age structure: below 43 degrees, avg age ~10, above 45 degrees, avg age ~15

ggplot(yt_n_survey_bio) +
  geom_point(aes(x = Age_years, y = Length_cm, col = Sex), alpha = 0.1)
# Large fish female, old fish male prevails, but not as notable as with canary.

yt_n_survey_bio |>
  group_by(Age_years) |>
  summarise(pct_f = sum(Sex == 'F')/n(),
            se = sqrt(pct_f * (1-pct_f) / n())) |>
  ggplot() +
  geom_point(aes(x = Age_years, y = pct_f)) +
  geom_errorbar(aes(x = Age_years, ymin = pct_f - se, ymax = pct_f + se))

```

